Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-09-05 papers/thesis-kirstie/thesis.tex
more cleanups

Breaking the Hamiltonians H into its component energies
\begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
where the translational kinetic energy of the N atoms is given by \begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
the potential energy of the total interaction of each atom with the other N-1 atoms is given by
\begin{displaymath}{V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} and \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} is an externally applied potential that varies spatially, such that its effects on each atom depend on the atom's position. Often $V_{ext}$ is set to zero, as it is in this paper.

Noting that the kinetic energy depends only on momentum while the total interatomic and external potential energies depend only on position, the integral in equation (8) can be separated into two parts 
\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume satisfying \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation} 
%the equation can be written
%\begin{equation}{Z=\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}e^\frac{-(\frac{|\vec{p}_2|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)\left(\frac{1}{V^N}\int{...}\int{dQ}e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}\right)}\end{equation}  
and the partitian function can be expressed as
\begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
where
\begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} The ideal partitian function $Z_{ideal}$ is associated with an ideal gas which exhibits no interaction between the atoms - a condition approached by a gas at low density where the atoms rarely encounter one another. The excess partitian function $Z_{excess}$ accounts for the interaction of all the atoms according to their interatomic potential. The free energy can also be broken into two parts using Equation~\ref{Zmultiply} giving
%\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})}\end{equation}
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}

@@

\begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
These integrals can conveniently be written in terms of the DeBroglie wavelength $\Lambda$
\begin{equation}{\frac{1}{h^{3}}\int{d\vec{p}}~e^\frac{-\frac{|\vec{p}|^2}{2m}}{k_BT}=\left(\sqrt{\frac{k_BTm}{2\pi\hbar^2}}\right)^3=\frac{1}{\Lambda^{3}}}\end{equation} 
where \begin{equation}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{equation} 
Rewriting Equation~\ref{Fidealmultipleint} in terms of $\Lambda$ gives
\begin{equation}{F_{ideal}= -k_BT[\ln(V^N)-\ln(N!) - \ln(\Lambda^{3N})]}\end{equation}Using Stirling's approximation \begin{displaymath}{\ln(N!)=N\ln{N}-N}\end{displaymath} this becomes
\begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} 
Multiplying by $\frac{V}{V}$ gives an expression in terms of the number density $n=\frac{N}{V}$ gives \begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N\Lambda^3}{V}\right)}-\frac{N}{V}\right]}V\end{equation}
\begin{equation}{F_{ideal}= k_BT[n\ln(n\Lambda^3)-n]V}\label{eq:Fideal}\end{equation}   
For a spatially varying number density, the ideal free energy in Equation~\ref{eq:Fideal} becomes
\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]d\vec{r}}\end{equation}

Now consider the excess free energy which is written as
\begin{equation}\label{Fexcess}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation} 
This integral can be greatly simplified if it can be assumed that the potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the potential energies between pairs of atoms, where each atom is paired in turn with each of the other atoms in the system. In this case $V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ becomes 
\begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
%\begin{displaymath}{=V_{12}(\vec{r_1},\vec{r_2})+V_{13}(\vec{r_1},\vec{r_3}) + V_{14}(\vec{r_1},\vec{r_4}) + ...}\end{displaymath}
%\begin{displaymath}{+ V_{23}(\vec{r_2},\vec{r_3})+V_{24}(\vec{r_2},\vec{r_4})+V_{25}(\vec{r_2},\vec{r_5})+ ...}\end{displaymath} 
%\begin{displaymath}{+ V_{34}(\vec{r_3},\vec{r_4})+V_{35}(\vec{r_3},\vec{r_5})+V_{36}(\vec{r_3},\vec{r_6})+...}\end{displaymath}
Using this potential, and setting the external potential $V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in equation \ref{Fexcess} becomes

>>>>

Breaking the Hamiltonian, H into its component energies
\begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
where the translational kinetic energy of the N atoms is given by \begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
the potential energy of the total interaction of each atom with the other N-1 atoms is given by
\begin{displaymath}{V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} and \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} is an externally applied potential that varies spatially, such that its effects on each atom depend on the atom's position. Often $V_{ext}$ is set to zero, as it is in this paper.

Noting that the kinetic energy depends only on momentum while the total interatomic and external potential energies depend only on position, the integral in equation (8) can be separated into two parts 
\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume satisfying \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation} 
%the equation can be written
%\begin{equation}{Z=\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}e^\frac{-(\frac{|\vec{p}_2|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)\left(\frac{1}{V^N}\int{...}\int{dQ}e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}\right)}\end{equation}  
the partitian function can be expressed as
\begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
where
\begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} The ideal partitian function $Z_{ideal}$ is associated with an ideal gas which exhibits no interaction between the atoms - a condition approached by a gas at low density where the atoms rarely encounter one another. The excess partitian function $Z_{excess}$ accounts for the interaction of all the atoms according to their interatomic potential. The free energy can also be broken into two parts using Equation~\ref{Zmultiply} giving
%\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})}\end{equation}
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}

@@

\begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(V^N\frac{1}{N!}\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
These integrals can conveniently be written in terms of the DeBroglie wavelength $\Lambda$
\begin{equation}{\frac{1}{h^{3}}\int{d\vec{p}}~e^\frac{-\frac{|\vec{p}|^2}{2m}}{k_BT}=\left(\sqrt{\frac{k_BTm}{2\pi\hbar^2}}\right)^3=\frac{1}{\Lambda^{3}}}\end{equation} 
where \begin{equation}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{equation} 
Rewriting Equation~\ref{Fidealmultipleint} in terms of $\Lambda$ gives
\begin{equation}{F_{ideal}= -k_BT[\ln(V^N)-\ln(N!) - \ln(\Lambda^{3N})]}\end{equation}Using Stirling's approximation \begin{displaymath}{\ln(N!)=N\ln{N}-N}\end{displaymath} this becomes
\begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} 
Rewriting the equation in terms of the number density $n=\frac{N}{V}$, and the quantum volume, $n_Q=\frac{1}{\lambda^3}$ gives
\begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N}{V}\frac{1}{n_Q}\right)}-\frac{N}{V}\right]}V\end{equation}
\begin{equation}{F_{ideal}= k_BT[n\ln(n/n_Q)-n]V}\label{eq:Fideal}\end{equation}   
For a spatially varying number density, the ideal free energy in Equation~\ref{eq:Fideal} becomes
\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})/n_Q)-n(\vec{r})]d\vec{r}}\end{equation}

Now consider the excess free energy which is written most generally as
\begin{equation}\label{Fexcess}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation} 
This integral can be greatly simplified if it can be assumed that the potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the potential energies between pairs of atoms, where each atom is paired in turn with each of the other atoms in the system. In this case $V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ becomes 
\begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
%\begin{displaymath}{=V_{12}(\vec{r_1},\vec{r_2})+V_{13}(\vec{r_1},\vec{r_3}) + V_{14}(\vec{r_1},\vec{r_4}) + ...}\end{displaymath}
%\begin{displaymath}{+ V_{23}(\vec{r_2},\vec{r_3})+V_{24}(\vec{r_2},\vec{r_4})+V_{25}(\vec{r_2},\vec{r_5})+ ...}\end{displaymath} 
%\begin{displaymath}{+ V_{34}(\vec{r_3},\vec{r_4})+V_{35}(\vec{r_3},\vec{r_5})+V_{36}(\vec{r_3},\vec{r_6})+...}\end{displaymath}
Using this potential, and setting the external potential $V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in Equation~\ref{Fexcess} becomes
